An application of Graph Theory to MDAO problem formulation

### 2012-10-02 intro.tex
fixed eqn's
updating the simple DSM

\section{Fundamental Problem Formulation}
    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem 
    the problem formulation is as follows:

    \begin{align}
        given & \ \ A: {x,y} \rightarrow {z}
              & \ \ B: {x,z} \rightarrow {y^t}
        min. &\ \ f(m,t) \notag
        w.r.t. & \ \ x,y,z \notag
        s.t. & \ \ G(y^t,y) = 0
        \label{eqn:simple}
    \end{align}

@@

    \begin{align}
        given & \ \ B: {x,z} \rightarrow {y}
              & \ \ A: {x,y} \rightarrow {z^t}
        min. &\ \ f(m,t) \notag
        w.r.t. & \ \ x,y,z \notag
        s.t. & \ \ G(z^t,z) = 0
        \label{eqn:simple2}
    \end{align}

@@

    \begin{align}
        given & \ \ A: {x,y} \rightarrow {z^t}
              & \ \ B: {x,z} \rightarrow {y^t}
        min. &\ \ f(m,t) \notag
        w.r.t. & \ \ x,y,z \notag
        s.t. & \ \ G1(z^t,z) = 0 \notag
             & \ \ G2(y^t,y) = 0
        \label{eqn:simple_fpf}
    \end{align}

@@

    \section{Problem Formulation Syntax}

    The same mathematical syntax can be used to describe more complex problem formulations
    such as ones that would decompose the FPF to a set of sub-problems. Tedford and Martins used this format to specify the 
    FPF for a set of test problems and also to describe specific formulations for solving them with a 
    number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
    problem formulations can all relate back to a common FPF.

    In the above problem formulation, there is no explicit relationship definition of local and global features. Instead they are 
    implicitly defined through variable naming conventions. For instance, $x$ is a global variable because it is used in $A.x$ and $B.x$. 
    Simliarly, $A.y$ is a local variable because it only affects the of $A$. If such a naming convention is not in place, 
    then a more explicit definition of the relation between variables must also be given in the form of additional equality constraints between 
    input variables. Similarly, both the constraint, $C$ and the residual, $R$ are presented as equality constraints. They could be classified as 
    a global and a local constraint on the problem respectively, with the only difference being the number of analysis tools that affect them.  
    Hence, the distinction between local and global properties are not necessary to form a fundamental 
    problem formulation. On the other hand, they are relevant when deriving a specific solution path for a problem, so any definition of problem 
    formulation should include the capability to detect the weather a given property is local or global.

    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed. 
    A number of matrix based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a DSM for the above notional example. 
    The DSM has been successfully applied in a number of ways to algorithmically modify problem formulations to 
    improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
    DSM  minimizing the computational costs of solving highly coupled systems\cite{Rogers1996}. Notice however, that the DSM in 
    Figure \ref{fig:dsm_simple} contains less information than the full problem formulation. Objective and constraint information 
    is missing and it's not possible to distinguish individual variables.

    \begin{figure}[!hbp]
        \begin{center}
        \includegraphics[width=.75\textwidth]{images/dsm_simple}
        \caption{Graph and matrix based DSM specifications for a notional analysis \label{fig:dsm_simple}}
        \end{center}
    \end{figure}

>>>>

\subsection{Fundamental Problem Formulation}
    Problem formulation is most commonly given in a mathematical syntax. For a simple, notional problem 
    the problem formulation is as follows:

    \begin{align}
        given & \ \ A: {x,y} \rightarrow {z}
        \\      & \ \ B: {x,z} \rightarrow {y^t}
        \\min. &\ \ f(m,t) \notag
        \\w.r.t. & \ \ x,y,z \notag
        \\s.t. & \ \ G(y^t,y) = 0
        \label{eqn:simple}
    \end{align}

@@

    \begin{align}
        given & \ \ B: {x,z} \rightarrow {y}
        \\      & \ \ A: {x,y} \rightarrow {z^t}
        \\min. &\ \ f(m,t) \notag
        \\w.r.t. & \ \ x,y,z \notag
        \\s.t. & \ \ G(z^t,z) = 0
        \label{eqn:simple2}
    \end{align}

@@

    \begin{align}
        given & \ \ A: {x,y} \rightarrow {z^t}
        \\      & \ \ B: {x,z} \rightarrow {y^t}
        \\min. &\ \ f(m,t) \notag
        \\w.r.t. & \ \ x,y,z \notag
        \\s.t. & \ \ G1(z^t,z) = 0 \notag
        \\     & \ \ G2(y^t,y) = 0
        \label{eqn:simple_fpf}
    \end{align}

@@

    \subsection{Problem Formulation Syntax}

    As shown above, the mathematical language for specifying problem formulations is very general and can be used both for 
    fundamental and specific problem formulations. Tedford and Martins used the above mathematical syntax to specify the 
    FPF for a set of test problems and also to describe specific formulations for solving them with a 
    number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
    problem formulations can all relate back to a common FPF.

    \begin{comment}
    In the above problem formulation, there is no explicit relationship definition of local and global features. Instead they are 
    implicitly defined through variable naming conventions. For instance, $x$ is a global variable because it is used in $A.x$ and $B.x$. 
    Simliarly, $A.y$ is a local variable because it only affects the of $A$. If such a naming convention is not in place, 
    then a more explicit definition of the relation between variables must also be given in the form of additional equality constraints between 
    input variables. Similarly, both the constraint, $C$ and the residual, $R$ are presented as equality constraints. They could be classified as 
    a global and a local constraint on the problem respectively, with the only difference being the number of analysis tools that affect them.  
    Hence, the distinction between local and global properties are not necessary to form a fundamental 
    problem formulation. On the other hand, they are relevant when deriving a specific solution path for a problem, so any definition of problem 
    formulation should include the capability to detect the weather a given property is local or global. 
    \end{comment}

    The challenge with using this traditional mathematical syntax is that it is not easily manipulated or analyzed. 
    A number of matrix based methods have been used successfully to translate the mathematical syntax into a more useful computational form. 
    Steward's Design Structure Matrix (DSM) is a square adjacency matrix which captures the relationship between analysis tools where off 
    diagonal elements of the matrix indicate coupling\cite{Steward1981}. Figure \ref{fig:dsm_simple} shows a DSM for the above notional example. 
    The DSM has been successfully applied in a number of ways to algorithmically modify problem formulations to 
    improve computational efficiency. Rogers et. al developed DeMAID to manipulate a
    DSM  to find an ordering for analysis tools that reduces the cost of solving highly coupled systems\cite{Rogers1996}. This re-ordering yields
    multiple specific problem formulations which all solve the same FPF. In other words, manipulation of the DSM does not fundamentally alter
    the problem formulation, which makes DSM an excellent foundation for specifying the FPF itself. Notice however, that the DSM in 
    Figure \ref{fig:dsm_simple} contains less information than the Eqn. \ref{eqn:simple_fpf}. Objective and constraint information 
    is missing and it's not possible to distinguish individual variables.

    \begin{figure}[!hbp]
        \begin{center}
        \includegraphics[width=.75\textwidth]{images/dsm_simple}
        \caption{Graph and matrix based DSM specifications for Eqn. \ref{eqn:simple_FPF} \label{fig:dsm_simple}}
        \end{center}
    \end{figure}
